---
title: "Assignment 10"
author: "Irene Jacob"
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(janeaustenr)
library(dplyr)
library(stringr)
library(tidytext)
library(tidyr)
library(ggplot2)
library(textdata)
library(wordcloud)
library(lexicon)
library(reshape2)
library(harrypotter)
```

# Goal

In Text Mining with R, Chapter 2 looks at Sentiment Analysis.  In this assignment, you should start by getting the primary example code from chapter 2 working in an R Markdown document.  You should provide a citation to this base code.  You’re then asked to extend the code in two ways:

- Work with a different corpus of your choosing, and
- Incorporate at least one additional sentiment lexicon (possibly from another R package that you’ve found through research).

As usual, please submit links to both an .Rmd file posted in your GitHub repository and to your code on rpubs.com.

# 1. Base Code from Text

### Citation:

Title: “Text Mining with R - A Tidy Approach”

Authors: Juilia Silge and David Robinson

Chapters: Chapter 2: [Sentiment Analysis with Tidy Data](https://www.tidytextmining.com/sentiment.html)

Date of publication: 2017

## Line Number and chapter

```{r}

tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, 
                                     regex("^chapter [\\divxlc]", 
                                                 ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)


nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)

jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)


ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

## Compare the 3 sentiment dictionaries

```{r}

pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

pride_prejudice

afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(pride_prejudice %>% 
                            inner_join(get_sentiments("bing")) %>%
                            mutate(method = "Bing et al."),
                          pride_prejudice %>% 
                            inner_join(get_sentiments("nrc") %>% 
                                         filter(sentiment %in% c("positive", 
                                                                 "negative"))) %>%
                            mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)

get_sentiments("nrc") %>% 
     filter(sentiment %in% c("positive", 
                             "negative")) %>% 
  count(sentiment)

#comparison graph
bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")
```

```{r}
get_sentiments("nrc") %>% 
     filter(sentiment %in% c("positive", 
                             "negative")) %>% 
  count(sentiment)

get_sentiments("bing") %>% 
  count(sentiment)
```

## Most common positive and negative words

```{r}
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

bing_word_counts

# Graph
bing_word_counts %>%
  group_by(sentiment) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(y = "Contribution to sentiment",
       x = NULL) +
  coord_flip()
```

## Stop words

```{r}
custom_stop_words <- bind_rows(tibble(word = c("miss"), 
                                          lexicon = c("custom")), 
                               stop_words)

custom_stop_words
```

## Wordclouds

```{r}
custom_stop_words <- bind_rows(tibble(word = c("miss"), 
                                          lexicon = c("custom")), 
                               stop_words)

custom_stop_words
```

## Negative and positive wordclouds

```{r}
tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("gray20", "gray80"),
                   max.words = 100)

```

## Looking at units beyond just words

```{r}
PandP_sentences <- tibble(text = prideprejudice) %>% 
  unnest_tokens(sentence, text, token = "sentences")

PandP_sentences$sentence[2]

austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text, token = "regex", 
                pattern = "Chapter|CHAPTER [\\dIVXLC]") %>%
  ungroup()

austen_chapters %>% 
  group_by(book) %>% 
  summarise(chapters = n())

bingnegative <- get_sentiments("bing") %>% 
  filter(sentiment == "negative")

wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
  summarize(words = n())

tidy_books %>%
  semi_join(bingnegative) %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("book", "chapter")) %>%
  mutate(ratio = negativewords/words) %>%
  filter(chapter != 0) %>%
  top_n(1) %>%
  ungroup()
```

# 2. Different Corpus of my choosing

I am quit fond of Harry Potter so wanted to see if there is some way I could work on it and found a package on it [**here**](https://github.com/bradleyboehmke/harrypotter) so i will be using it.

The below code is based on this [*link*](https://uc-r.github.io/sentiment_analysis)

## Chapter 1 of Goblet of Fire
```{r}
# lets consider 1 chapter from *goblet_of_fire*

goblet_of_fire[1:1]
```

## Tidying the data

The chapters are tokenised into words and the head and tail of the data frame are displayed.

```{r}
titles <- c("Goblet of Fire")

books <- list(goblet_of_fire)
  
series <- tibble()

for(i in seq_along(titles)) {
        
        clean <- tibble(chapter = seq_along(books[[i]]),
                        text = books[[i]]) %>%
             unnest_tokens(word, text) %>%        #tokenize each chapter to words
             mutate(book = titles[i]) %>%
             select(book, everything())

        series <- rbind(series, clean)
}

# set factor to keep books in order of publication
series$book <- factor(series$book, levels = rev(titles))

head(series)
tail(series)
```

## Different sentiments in this dataset using loughran

```{r}
series %>%
        right_join(get_sentiments("loughran")) %>%
        filter(!is.na(sentiment)) %>%
        count(sentiment, sort = TRUE)
```

## Visualize

```{r}
series %>%
        group_by(book) %>% 
        mutate(word_count = 1:n(),
               index = word_count %/% 500 + 1) %>% 
        inner_join(get_sentiments("bing")) %>%
        count(book, index = index , sentiment) %>%
        ungroup() %>%
        spread(sentiment, n, fill = 0) %>%
        mutate(sentiment = positive - negative,
               book = factor(book, levels = titles)) %>%
        ggplot(aes(index, sentiment, fill = book)) +
          geom_bar(alpha = 0.5, stat = "identity", show.legend = FALSE)
```

## Comparison between AFINN and LOU

```{r}
afinn <- series %>%
        group_by(book) %>% 
        mutate(word_count = 1:n(),
               index = word_count %/% 500 + 1) %>% 
        inner_join(get_sentiments("afinn")) %>%
        group_by(book, index) %>%
        summarise(sentiment = sum(value)) %>%
        mutate(method = "AFINN")

loughran <- series %>%
            group_by(book) %>% 
            mutate(word_count = 1:n(),
                  index = word_count %/% 500 + 1) %>% 
            inner_join(get_sentiments("loughran")) %>%
            mutate(method = "loughran") %>%
count(book, method, index = index , sentiment) %>%
        ungroup() %>%
        spread(sentiment, n, fill = 0) %>%
        mutate(sentiment = positive - negative) %>%
        select(book, index, method, sentiment)

```

```{r}
bind_rows(loughran, afinn) %>%
        ungroup() %>%
        mutate(book = factor(book, levels = titles)) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_bar(alpha = 0.8, stat = "identity", show.legend = FALSE) +
  facet_grid(book ~ method)
```

# 3. New sentiment lexicon - Loughran

I found Loughran Sentiment Lexicon [*here*](https://www.rdocumentation.org/packages/tidytext/versions/0.2.6/topics/get_sentiments).

```{r}
get_sentiments("loughran")
```