---
title: "Assignment 9"
author: "Irene Jacob"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(XML)
library(RCurl)
library(jsonlite)
library(DT)
library(tidyverse)
```

## Goal 

The New York Times web site provides a rich set of APIs, as described [**here**](https://developer.nytimes.com/apis). You’ll need to start by signing up for an API key. Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R DataFrame.

## Register for API Key and Retrieve the Data

Read the data having movie reviews of Harry Potter, GOdfather and Jurassic Park from the API.

```{r}
nyt_key <- "VnkQnIsatF3PkvjPsr78YUYFUtkfIvdz"
url_1 <- "https://api.nytimes.com/svc/movies/v2/reviews/search.json?query=harry%20potter&api-key="

data_1 <- fromJSON (paste0(url_1,nyt_key))

url_2 <- "https://api.nytimes.com/svc/movies/v2/reviews/search.json?query=godfather&api-key="

data_2 <- fromJSON (paste0(url_2,nyt_key))

url_3 <- "https://api.nytimes.com/svc/movies/v2/reviews/search.json?query=jurassic%20park&api-key="

data_3 <- fromJSON (paste0(url_3,nyt_key))
```

## Create DataFrame

Load into 3 data frames

```{r}
data_1 <- data.frame(data_1)
data_2 <- data.frame(data_2)
data_3 <- data.frame(data_3)
```

## Tidy the data

The 3 dataframes are combined into a single data frame 'final'.
The required coloumns are selected from the dataframe and the columns are renamed.

```{r}
final <- dplyr::bind_rows(data_1,data_2,data_3)

final <- final %>% select(results.display_title, results.byline, results.headline, results.summary_short, results.publication_date)

names <- c('Title', 'Author', 'Headline', 'Short Summary', 'Date of Publication')

colnames(final) <- names

datatable(final)

```

## Conclusion

The data was read successfully from API and loaded into dataframes. 
Necessary tidying was done to the 'final' dataframe.