---
title: "Assignment 3"
author: "Irene Jacob"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(openintro)
library(htmlwidgets)

```

## 1. 
Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

#### Load the data from csv file and identify the majors that contain DATA and STATISTICS

```{r eval = TRUE}
majors <- read.csv("https://raw.githubusercontent.com/datasets/five-thirty-eight-datasets/master/datasets/college-majors/data/majors-list.csv")

temp <- majors$major

dt<-grepl("data", temp)       #identify majors containing DATA

temp[dt]
```

```{r eval = TRUE}
st<-grepl("statistics", temp)   #identify majors containing STATISTICS

temp[st]

```

\clearpage
## 2.

Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r eval = TRUE, message = FALSE}
s1 = c("bell pepper", "bilberry", "blackberry","blood orange")
s2 = c("blueberry","cantalope","chili pepper","cloudberry")
s3 = c("elderberry","lime","lychee","mulberry")
s4 = c("olive","salal berry")

berries <- c(s1,s2,s3,s4)
print(berries)

```

\clearpage
## 3.
Describe, in words, what these expressions will match:

##### 1. (.)\1\1

A character will be repeated thrice. Example iii.

```{r eval = TRUE}
l <- c("abba","cac","aabb","absca","abcdefg","aaaaaa","iii","abab","i1221i","i123456721i","mpmnm","0101","acca")
str_view(l, "(.)\\1\\1", match = TRUE)
```

##### 2. "(.)(.)\\2\\1"

One pair of characters will be repeated twice but the second time will be the reverse of the first. Example acca.

```{r eval = TRUE}
str_view(l, "(.)(.)\\2\\1", match = TRUE)
```

##### 3. (..)\1

2 characters will be repeated. It can be any 2 characters. Example 0101.

```{r eval = TRUE}
str_view(l, "(..)\\1", match = TRUE)
```

##### 4. "(.).\\1.\\1"

Example mpmnm. A character followed by another character then the first character again, then a new character followed by the first character again.

```{r eval = TRUE}
str_view(l, "(.).\\1.\\1", match = TRUE)
```

##### 5. "(.)(.)(.).*\\3\\2\\1"

In the beginning there are 3 characters and the same 3 characters are there in the end of the string but in the reverse order. In the middle there are zero or more characters.
Example i123456721i or i1221i

```{r eval = TRUE}
str_view(l, "(.)(.)(.).*\\3\\2\\1", match = TRUE)
```

\clearpage
## 4.
Construct regular expressions to match words that:

##### 1. Start and end with the same character. 

"^(.).*\\1$"

```{r eval = TRUE}
str_view(l, "^(.).*\\1$", match = TRUE)
```

##### 2. Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

"(.)(.).*\\1\\2"

```{r eval = TRUE}
str_view(berries, "(.)(.).*\\1\\2", match = TRUE)
```

##### 3. Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

"(.).*\\1.*\\1"

```{r eval = TRUE}
str_view(berries, "(.).*\\1.*\\1", match = TRUE)
```