---
title: "Project 2"
author: "Irene Jacob"
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(DT)
library(tidyverse)
library(openintro)
library(countrycode)
library(ggplot2)
```

## Goal
The goal of this assignment is to give you practice in preparing different datasets for downstream analysis work.

Your task is to:

Choose any three of the “wide” datasets identified in the Week 6 Discussion items. (You may use your own dataset; please don’t use my Sample Post dataset, since that was used in your Week 6 assignment!) For each of the three chosen datasets:

 Create a .CSV file (or optionally, a MySQL database!) that includes all of the information included in the dataset. You’re encouraged to use a “wide” structure similar to how the information appears in the discussion item, so that you can practice tidying and transformations as described below.

 Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data. [Most of your grade will be based on this step!]

 Perform the analysis requested in the discussion item.

 Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative descriptions of your data cleanup work, analysis, and conclusions.

## Dataset 1
### read the data

The data is read from a .csv file located in my github. It can be accessed [**here**](https://github.com/irene908/DATA-607/blob/master/dataset1_project2.csv)

The column names had a X before the name so I used some regular expression to clean that out.
```{r}
data1 <- read.csv("https://raw.githubusercontent.com/irene908/DATA-607/master/dataset1_project2.csv")
names(data1)[1] <- "Country"
names(data1) <- gsub(x = names(data1), pattern = "\\X", replacement = "")
datatable(data1, options = list(scrollY = 300, scroller = TRUE ))
nrow(data1)
```

### remove the rows which are not countries

```{r}
excess<- c('North America', 'Central & South America', 'Antarctica', 'Eurasia', 'Middle East', 'Asia & Oceania', 'World', 'Africa', 'Europe', 'Former Czechoslovakia', 'Former Serbia and Montenegro', 'Former Yugoslavia', 'East', 'Hawaiian Trade Zone', 'U.S. Pacific Islands', 'Wake Island', 'Former U.S.S.R.')

data1_new <- data1[ !grepl(paste(excess, collapse="|"), data1$Country),]

datatable(data1_new, options = list(scrollY = 300, scroller = TRUE ))
nrow(data1_new)
```

### Create a list of country code

A separate list of the country codes of each of the countries is created and it will be appended to the original data later on.

```{r}
Continent <- as.factor(countrycode(sourcevar = data1_new[, "Country"], origin = "country.name", destination = "continent"))
Continent<-data.frame(Continent)
datatable(Continent, options = list(scrollY = 300, scroller = TRUE ))
nrow(Continent)
```

### Append data and the country code list

```{r}
data1_append<-data.frame(Continent, data1_new)
names(data1_append) <- gsub(x = names(data1_append), pattern = "\\X", replacement = "")
datatable(data1_append, options = list(scrollY = 300, scroller = TRUE ))
nrow(data1_append)
```

### convert to long 

The tidyr gather() function is used.
```{r}
data1_long <- df %>% gather(data=data1_append, Population, "1980":"2010")
colnames(data1_long)[colnames(data1_long)=="."]<-"Year" 

datatable(data1_long, options = list(scrollY = 300, scroller = TRUE ))
nrow(data1_long) 
```

### Analyse the change in population in FRANCE

```{r}
data1_france<-subset(data1_long, Country=='France', select=c(Country, Year, Population))

ggplot(data=data1_france, aes(x=Year, y=Population, group=1)) +
  geom_line()+
  geom_point()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Analyse the change in population in Martinique

```{r}
data1_Martinique <-subset(data1_long, Country=='Martinique', select=c(Country, Year, Population))
ggplot(data=data1_Martinique, aes(x=Year, y=Population, group=1)) +
  geom_line()+
  geom_point()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Dataset 2
### read the property crime data

The data is read from a .csv file located in my github. It can be accessed [**here**](https://github.com/irene908/DATA-607/blob/master/dataset2_project2.csv)

```{r}
data2_property <- read.csv("https://raw.githubusercontent.com/irene908/DATA-607/master/dataset2_project2.csv")
names(data2_property) <- gsub(x = names(data2_property), pattern = "\\X", replacement = "")
datatable(data2_property, options = list(scrollY = 300, scroller = TRUE ))
```

### convert to long 

```{r}
data2_property_long <- data2_property %>% 
  select(-16) %>%                   #omitting the last empty column
  gather("Year", "Total", 3:15)
datatable(data2_property_long, options = list(scrollY = 300, scroller = TRUE ))
```

### read the violent crime data

The data is read from a .csv file located in my github. It can be accessed [**here**](https://github.com/irene908/DATA-607/blob/master/dataset2.1_project2.csv)

```{r}
data2_violent <- read.csv("https://raw.githubusercontent.com/irene908/DATA-607/master/dataset2.1_project2.csv")
names(data2_violent) <- gsub(x = names(data2_violent), pattern = "\\X", replacement = "")
datatable(data2_violent, options = list(scrollY = 300, scroller = TRUE ))
```

### convert to long 

```{r}
data2_violent_long <- data2_violent %>% 
  select(-16) %>%                   #omitting the last empty column
  gather("Year", "Total", 3:15)
datatable(data2_violent_long, options = list(scrollY = 300, scroller = TRUE ))
```

### Combine the 2 dataframes

Using a for loop the two dataframes are merged together

```{r}
data2_violent_long$Crime <- "Violent"
data2_property_long$Crime <- "Property"

data2_crime = data.frame(matrix(vector(), 286, 5,
                dimnames=list(c(), c("Agency", "State", "Year", "Total", "Crime"))))

for(i in 1:5){
data2_crime[i] <- bind_rows(data2_violent_long[i], data2_property_long[i])
}

datatable(data2_crime, options = list(scrollX = TRUE))
```

### Analysis

The crime rate is low for NEW YORK CITY and the change is also low for this city.

```{r}
ggplot(data2_crime, aes(Year, Total)) + geom_bar(stat = "identity", position=position_dodge(), aes(fill = Crime)) + facet_wrap(.~Agency) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Crime rate in all the cities

```{r}
#crime rates for all cities over time
ggplot(data2_crime) + 
  geom_point(aes(Year, Total, color=Agency))
```

## Dataset 3
### read the data

The data is read from a .csv file located in my github. It can be accessed [**here**](https://github.com/irene908/DATA-607/blob/master/dataset3_project2.csv)

```{r}
data3 <- read.csv("https://raw.githubusercontent.com/irene908/DATA-607/master/dataset3_project2.csv")
names(data3)[1] <- "Year"
names(data3) <- gsub(x = names(data3), pattern = "\\X", replacement = "")
data3 <- data3[-c(1), ]
datatable(data3, options = list(scrollY = 300, scroller = TRUE ))
```

### cleaning the data and creating a subset of the required data

```{r}
data3_new<-rename(data3,ARG=Argentina,SWE=Sweden,US=United.States,KOR=Korea..Rep.,JPN=Japan,BRA=Brazil,CAN=Canada,IND=India,MEX=Mexico,IRE=Ireland)
data3_new<-select(data3_new,Year,ARG,SWE,US,KOR,JPN,BRA,CAN,IND,MEX,IRE)
datatable(data3_new, options = list(scrollY = 300, scroller = TRUE ))
```

### Convert to long

```{r}
data3_long <- gather(data3_new,Country,Rate_Unemployment,-Year)
datatable(data3_long, options = list(scrollY = 300, scroller = TRUE ))
```

### Analysis of unemployment in IRELAND

```{r}
data3_long_ire<-subset(data3_long,Country=="IRE")
ggplot(data3_long_ire, aes(x=Year, y=Rate_Unemployment)) + geom_bar(stat='identity')
```

### Analysis of unemployment in INDIA

```{r}
data3_long_ind<-subset(data3_long,Country=="IND")
ggplot(data3_long_ind, aes(x=Year, y=Rate_Unemployment)) + geom_bar(stat='identity')
```